On the minimum leaf conductance: its role in models of plant water use, and ecological and environmental controls

### 2017-09-19 manuscript.Rmd
Switched to tansley review format, lots of reorganizing

---
title: "On the minimum leaf conductance: ecological and environmental controls, and its role in models of plant water use"
author: ''
date: ''
output:
  word_document:
    reference_docx: manuscriptstyle.docx
  html_document:
    number_sections: yes
    theme: flatly
csl: global-change-biology.csl
bibliography: references.bib
---

@@

# Abstract

@@

# Introduction

\
Stomata exert a strong control on the transpiration rate of leaves [@bange1953]. It is well known that stomata respond quickly to changes in light intensity, humidity and CO~2~ concentration at the leaf surface. In doing so, stomata can be thought to balance the uptake of CO~2~ with the loss of water, and thus can be expected to close stomata when photosynthetic rates are low. To avoid desiccation and ultimate death, stomata can also be expected to close during periods of severe water stress. Plants cannot, however, completely halt water loss even when stomata are closed. In this review, we discuss various aspects of the minimum leaf conductance reached when stomata are essentially closed. We pay special attention to the role that minimum conductance plays in land-surface models, in particular as it forms part of the much-used 'Ball-Berry'-type class of stomatal conductance models.

@@

\
Where g~1~ is a 'slope' parameter, A~n~ the leaf net photosynthetic rate, C~a~ atmospheric CO~2~ concentration, and f(D) is some function of the vapour pressure deficit D (or relative humidity in case of the model of @ballberry1987), and g~0~ is the value of g~s~ when A~n~ is zero. In this paper, we refer to Eq. 1 as 'the Ball-Berry model', thus including all model formulations that include different f(D) functions besides the one proposed by Ball & Berry (1987) (they used relative humidity).

\
The effects of photosynthetic photon flux density (PPFD) and leaf temperature are both assumed to enter via the dependence of A~n~ on these drivers. The g~0~ parameter is another definition of the minimum conductance, but we keep the notation g~0~ to indicate that it is a parameter in the Ball-Berry model. Much attention has been paid to quantifying and interpreting variation in the g~1~ parameter [@lin2015; @miner2017; @prentice2014; @medlyn2011], but the inocuous g~0~ parameter has been largely ignored. For example, a recent review of stomatal conductance models did not mention the minimum conductance [@damour2010]. The original description of Eq. 1 as published by @ballberry1987 did not include an intercept term. It was first introduced in the unpublished manuscript of @ball1988, as an 'intercept' without further discussion of the role it plays in the model.

\
The g~0~ parameter not only sets a minimum value of g~s~ in the model, it also modifies the behaviour of Eq. 1 after it has been coupled to the Farquhar-von Caemmerer-Berry model of photosynthesis (the so-called 'coupled leaf gas exchange model'). @leuning1990 defined g~0~ as g~s~ when A approaches zero as PPFD approaches the light compensation point, and pointed out that g~0~ is necessary to simulate the increase in A/g~s~ (and C~i~) at low light [see also @leuning1995]. Similarly, @collatz1992 showed that including g~0~ affects the response of g~s~ to relative humidity in the coupled leaf gas exchange model. Despite these early reports, the exact role of g~0~ in models is often overlooked.

@@

\
In this paper, we a) review the role and current use of g~0~ in models, b) perform a quantitative comparison of various measurements of the presumed minimum conductance, including fitting g~0~ with Eq. 1, night-time conductance (based on a previous compilation), and a new compilation of g~min~ estimates from weight-loss of detached leaves, and c) with use of two case studies and review of the literature, we demonstrate that g~min~ shows considerable acclimation to growing conditions.

# Methods

## Coupled leaf gas exchange model

To illustrate the effect of g~0~ on the modelled response of g~s~ to PPFD and T~leaf~, it is necessary to embed the stomatal conductance model in a coupled leaf gas exchange model, which is formed by three equations; Eq. 1 giving the dependence of g~s~ on A~n~ and some environmental drivers, the diffusion constraint,

## Literature review

For the @lin2015 database, we fit the model of @medlyn2011 using non-linear regression, while @miner2017 presented results from linear regression of the @ballberry1987 model.

\
Measurements of conductance of isolated, non-stomatal cuticles (g~cut,isol~), and of conductance of detached leaves with the stomatal side sealed (g~cut,seal~) were taken from the compilation by @kerstiens1996.

\
Data on minimum conductance (g~min~) from weight loss of detached leaves was compiled from 44 original papers, see Appendix A for a list of papers. For the compilation, we returned to all original papers mentioned in the review by @kerstiens1996, and added many newer sources. All conductance values were converted to per unit projected surface area. Units were converted to molar fluxes assuming 41 mol m^-3^ (cf. @kerstiens1996).

\
A compilation of night-time conductance (g~dark~) values was taken from @lombardozzi2017, who compiled cuvette (and sometimes sap flow-based) measurements of ca. 150 species during night-time. Actual minimum daytime values were estimated from the Lin et al. database, by taking those data where A~n~ < 2 $mu$ mol m^-2^ s^-1^, and calculating the minimum g~s~ for each of the XXX studies included in that dataset. This estimate was added as a reference as it represents an actual minimum values measured, rather than a fitted/extrapolated g~0~ term.

## Case studies on minimum conductance of detached leaves

Both studies use identical methods. The first dataset was collected on *Eucalyptus parramattensis* growing in twelve whole-tree chambers (see Drake for methods), either tracking ambient conditions or exposed to a constant +3 C air temperature. Leaves were sampled from XXX, n=XXX. The second case study was on *Hakea* plants. Details here (measurement T?). Methods for gmin from drying curves here.

# The use of g~0~ in models

@@

\
In the following we show how inclusion of g~0~ in models affects the response of C~i~ at low PPFD and the response of transpiration to high leaf temperature (T~leaf~). Another reason to include a non-zero g~0~ term in models of stomatal conductance arises during severe drought. When plants are sufficiently water-stressed that stomata are completely closed, water loss still continues through the cuticle and leaky and incompletely closed stomata. Thus models that aim to predict when plants desiccate and die must include a minimum conductance term [@blackman2016; @sinclair2000].

@@

# Estimation of g~0~ by fitting

# Comparison of various definitions of minimum conductance

| Variable   |  Definition |
| ---------- | ----------- |
| g~cut~     | Conductance of an isolated non-stomatal cuticle         |
| g~min~     | Minimum conductance measured from weight loss of detached leaves |
| g~dark~    | Night-time conductance                                  |
| g~0~       | Intercept in the Ball-Berry-type stomatal conductance model, fitted. |
| g~s,low A~ | Cuvette measurements of g~s~ when photosynthesis rates are low. |

Measurements of (presumed) minimum conductance, after stomatal closure is either induced or assumed, can be broadly divided into the following categories. The absolute minimum conductance attained by leaves is that through the cuticle only. Measurements of cuticular conductance (g~cut~) are typically made on isolated cuticles [@riederer2001] (g~cut,isol~), or on the side of the leaf that lacks stomata ('adaxial' side), or by sealing the side with stomata (g~cut,seal~). Another measurement is to detach leaves, and measure leaf weight over time as the leaf dries out. After some time, a constant minimum transpiration rate is attained (sometimes described as the 'residual transpiration rate'), from which minimum conductance (g~min~) can be inferred using measured vapour pressure deficit. This method most closely resembles conditions that plants may experience during a dry-down or periods of extreme stress. Because night-time conductance (g~dark~) is frequently used as an independent measurement of g~0~, we compare literature values of g~dark~ with g~min~ and g~cut~, as well as the compilation of fitted values of g~0~ (already shown in Fig. 3).

Results shown in Fig. 4.

\
Conductance in the dark was on average eight-fold higher than g~min~. It is no new observation that stomata do not close in the dark, but do close on detached leaves, as already demonstrated by @hygen1954 in Scots pine, and described by @koerner1994. More recently, @walden-coleman2013 showed much higher g~dark~ compared to g~min~ in Cotton.

\
Actual measured stomatal conductance during conditions when photosynthesis was low (g~s,low A~) was also much lower than fitted g~0~ or g~dark~. In other words, actual daytime conductance is much lower than the presumed minimum estimated from fitted g~0~. A similar result was already presented by @dekauwe2015 (their  Fig. A2). If we would use the fitted g~0~ in models, simulated g~s~ would never drop below g~0~.

@cavender-bares2007 gdark similar to gmin in oaks (unusual!).

### Low conductance of isolated cuticles

@@

Although we use the term cuticular conductance freely, this transport pathway does not represent a true conductance since water does not diffuse as a gas through the cuticle. Instead, it dissolves into the medium of the cuticle, diffuses through the solid matrix, and is desorped at the outer edge of the cuticle [@kerstiens1996; @riederer2001; @schreiber1996]. The main barrier to diffusion is actually a very thin layer at the leaf surface. Because most of the resistance is located in such a thin layer, g~cut~ does not correlate with the thickness of the cuticle  [@priestley1943; @riederer2001; @anfodillo2002; @schuster2016b]. Nonetheless there is considerable variation in cuticle thickness between plant species [@schuster2016a], as well as within species in response to drought (cite), along altitudinal transects [@delucia1984], and even with increasing height in the canopy of very tall trees [@woodruff2010]. If the cuticle thickness does not directly affect its conductance, what is the value of increasing the cuticle?

It is important to bear in mind other functions of the cuticle, including attenuation of radiation, as a barrier for fungal pathogens, and various interactions with insects [@mueller2008] [@kerstiens1996c;@riederer2008]. The thickness of the cuticle per se may also have to do with mechanical strength, as @onoda2012 reported that thicker cuticles are more resistant to tearing (the force to tear and tensile strength both approx proportional to cuticle thickness). Thick cuticles may also be useful because  exposure to wind has been shown to dislodge cuticular compounds by absrasion [@rogge1993], thereby dramatically increasing the conductance of the cuticle [@hadley1983;@grace1974;@vangardingen1991].

@grace1974: in a *Festuca* species, g~min~ more than doubled after exposure to high wind

Similarly, exposure to simulated rain damages the cuticle and increases its conductance [@baker1986]. Such exposure to the elements may be responsible for the typical gradual increase in g~min~ with leaf age in field-grown plants [@jordan2007].

Despite considerable work on the topic, no clear relationship between chemical composition or structure of the cuticle and its conductance has emerged. Recently, @schuster2016a presented a comprehensive study of cuticle chemical composition, but this variation could not explain variation in g~cut~. Similarly, @hauke1998 found no relationship between the gradual decrease in g~cut~ in *Hedera helix* during leaf maturation and cuticle wax amount, mean chain lengths, or cuticle weight.

Water travels more freely across the cuticle than CO2 [@boyer1997; @boyer2015]. This observation was used by @manzoni2011 to recalculate leaf water-use efficiency for droughted plants, but we caution against this because g~cut~ is only one component of g~min~, and CO~2~ will travel through leaky and incompletely closed stomata.

# Nighttime conductance

@@

# Minimum conductance of detached leaves

@@

@@

## Minimum conductance shows little meaningful interspecific variation

@@

Comparing all available data, we did not find significant relationships between rainfall, aridity, or temperature and g~min~. There is, however, a substantial acclimatory response of g~min~ to water availability.

## Minimum conductance acclimates to growing conditions

### Drought

@@

@@

### Altitude

### Leaf age

Also discussed above!
Bunch of papers in gmindatabase have age.

### Humidity and temperature

@@

Lower g~min~ result is consistent with @duarte2016, who reported lower g~0~ and g~dark~ in a heat wave treatment (Douglas-fir), which persisted for some time. This is not due to a direct effect of temperature on in cuticular conductance, which increases with temperature (Schreiber 2001). Our results showed a large  acclimation effect but no evidence for a direct T effect (over the narrow range in T compared to Schreiber).

@@

>>>>

---
title: "On the minimum leaf conductance: its role in models of plant water use, and ecological and environmental controls"
author: ''
date: ''
output:
  word_document:
    reference_docx: manuscriptstyle.docx
  html_document:
    number_sections: yes
    theme: flatly
csl: new-phytologist.csl
bibliography: references.bib
---

@@

# Summary

@@

# I. Introduction

\
Stomata exert a strong control on the transpiration rate of leaves [@bange1953]. It is well known that stomata respond quickly to changes in light intensity, humidity and CO~2~ concentration at the leaf surface. In doing so, stomata can be thought to balance the uptake of CO~2~ with the loss of water, and thus can be expected to close stomata when photosynthetic rates are low. To avoid desiccation and ultimate death, stomata can also be expected to close during periods of severe water stress. Plants cannot, however, completely halt water loss even when stomata are closed.

@@

\
Where g~1~ is a 'slope' parameter, A~n~ the leaf net photosynthetic rate, C~a~ atmospheric CO~2~ concentration, and f(D) is some function of the vapour pressure deficit D (or relative humidity in case of the model of @ballberry1987), and g~0~ is the value of g~s~ when A~n~ is zero. In this paper, we refer to Eq. 1 as 'the Ball-Berry model', thus including all model formulations that include different f(D) functions besides the one proposed by Ball & Berry (1987) (they used relative humidity). It should be clear from Eq. 1 that the Ball-Berry model is actually a model of water-use efficiency, since both g~s~ and A~n~ appear in the equation. Much attention has been paid to quantifying and interpreting variation in the g~1~ parameter [@lin2015; @miner2017; @prentice2014; @medlyn2011], but the g~0~ parameter has been largely ignored. For example, a recent review of stomatal conductance models did not mention the minimum conductance [@damour2010]. The original description of Eq. 1 as published by @ballberry1987 did not include an intercept term. It was first introduced in the unpublished manuscript of @ball1988, as an 'intercept' without further discussion of the role it plays in the model. Discuss @barnard2013 briefly (in a positive way :)

\
The absolute minimum attainable water loss rate through the cuticle only, which is typically measured on isolated cuticles of the adaxial (non-stomatal) side of the leaf. Other measurements allow for the estimation of water loss through incompletely closed and broken stomata. Of particular interest is the method where leaves are detached from the plant, and their weight loss monitored over time. This method (mass loss of detached leaves, MLD) is interesting because it more closely simulates actual field conditions for plants that are severely droughted, and includes all surfaces of the leaf. Since the seminal review by @kerstiens1996, no attempt has been made to synthesize the literature for values of g~min~, which is timely as models start to incorporate more detailed drought dynamics all the way to the point of mortality by desiccation (Xu, Blackman, Martin-StPaul, Sperry?). Instead, many models rely on inaccurately estimated 'minimum' conductance from cuvette data, by applying regression techniques. We critically examine this method in Section III.

\
Previews reviews have focused on the biology of the plant cuticle, and detailed physiology and anatomy of water transport across cuticles (riederer, kerstiens). Thus there is a disconnect between dizzying level of detailed understanding of the plant cuticle [@riederer2008], and the very simple  assumptions made when applying ecosystem models. We aim to improve the connection somewhat, and point at obvious gaps in understanding.

# II. Comparison of various definitions of minimum conductance

| Variable   |  Definition |
| ---------- | ----------- |
| g~cut~     | Conductance of an isolated non-stomatal cuticle.         |
| g~min~     | Minimum conductance measured from weight loss of detached leaves. |
| g~dark~    | Night-time conductance, or conductance after significant dark adaptation.              |
| g~0~       | Intercept in the Ball-Berry-type stomatal conductance model, i.e. g~s~ when A~n~ approaches zero. |
| g~s,low A~ | Cuvette measurements of g~s~ when photosynthesis rates are very low. |

Measurements of (presumed) minimum conductance, after stomatal closure is either induced or assumed, can be broadly divided into the following categories. The absolute minimum conductance attained by leaves is that through the cuticle only. Measurements of cuticular conductance (g~cut~) are typically made on isolated cuticles [@riederer2001] (g~cut,isol~), or on the side of the leaf that lacks stomata ('adaxial' side), or by sealing the side with stomata (g~cut,seal~). Another measurement is to detach leaves, and measure leaf weight over time as the leaf dries out. After some time, a constant minimum transpiration rate is attained (sometimes described as the 'residual transpiration rate'), from which minimum conductance (g~min~) can be inferred using measured vapour pressure deficit [@sinclair1986]. This method most closely resembles conditions that plants may experience during a dry-down or periods of extreme stress.

\
For the @lin2015 database, we fit the model of @medlyn2011 using non-linear regression, while @miner2017 presented results from linear regression of the @ballberry1987 model. Measurements of conductance of isolated, non-stomatal cuticles (g~cut,isol~), and of conductance of detached leaves with the stomatal side sealed (g~cut,seal~) were taken from the compilation by @kerstiens1996.  Data on minimum conductance (g~min~) from weight loss of detached leaves was compiled from 44+ original papers, see SuppInfo A for a list of papers and additional figures describing this new compilation. For the compilation, we returned to all original papers mentioned in the review by @kerstiens1996, and added many newer sources. All conductance values were converted to per unit projected surface area. Units were converted to molar fluxes assuming 41 mol m^-3^ (cf. @kerstiens1996). Although it is quite clear that night-time conductance (g~dark~) is much larger than a true 'minimum' conductance, it is frequently used as an independent measurement of g~0~ [@lombardozzi2017], and thus a comparison of magnitudes is useful. A compilation of night-time conductance (g~dark~) values was taken from @lombardozzi2017, who compiled cuvette (and sometimes sap flow-based) measurements of ca. 150 species during night-time. Actual minimum daytime values were estimated from the Lin et al. database, by taking those data where A~n~ < 2 $mu$ mol m^-2^ s^-1^, and calculating the minimum g~s~ for each of the XXX studies included in that dataset. This estimate was added as a reference as it represents an actual minimum values measured, rather than a fitted/extrapolated g~0~ term.

\
Conductance in the dark was on average eight-fold higher than g~min~. It is no new observation that stomata do not close in the dark, but do close on detached leaves, as already demonstrated by @hygen1954 in Scots pine, and described by @koerner1994. More recently, @walden-coleman2013 showed much higher g~dark~ compared to g~min~ in Cotton. However, @cavender-bares2007 gdark similar to gmin in oaks. Actual measured stomatal conductance during conditions when photosynthesis was low (g~s,low A~) was also much lower than fitted g~0~ or g~dark~. In other words, actual daytime conductance is much lower than the presumed minimum estimated from fitted g~0~. A similar result was already presented by @dekauwe2015 (their  Fig. A2).

# III. Estimating minimum conductance from cuvette data

# IV. Use of minimum conductance in models

## 1. Models of water-use efficiency

\
The effects of photosynthetic photon flux density (PPFD) and leaf temperature on g~s~ are both assumed to enter via the dependence of A~n~ on these drivers. The g~0~ parameter not only sets a minimum value of g~s~ in the model, it also modifies the behaviour of Eq. 1 after it has been coupled to the Farquhar-von Caemmerer-Berry model of photosynthesis (the so-called 'coupled leaf gas exchange model'). @leuning1990 defined g~0~ as g~s~ when A approaches zero as PPFD approaches the light compensation point, and pointed out that g~0~ is necessary to simulate the increase in A/g~s~ (and C~i~) at low light [see also @leuning1995]. Similarly, @collatz1992 showed that including g~0~ affects the response of g~s~ to relative humidity in the coupled leaf gas exchange model. Despite these early reports, the exact role of g~0~ in models is often overlooked.

@@

@@

@@

## 2. Models of drought response of stomatal conductance

Another reason to include a non-zero g~0~ term in models of stomatal conductance arises during severe drought. When plants are sufficiently water-stressed that stomata are completely closed, water loss still continues through the cuticle and leaky and incompletely closed stomata. Thus models that aim to predict when plants desiccate and die must include a minimum conductance term [@blackman2016; @sinclair2000].

@pisekwinkler1953, early calculation of the length of time needed to desiccate leaves to some water content, given the minimum transpiration rate, and the water content of the leaves.

@sinclair2000 gmin as a drought tolerance trait, can calculate time needed to death during drought phase when plants no longer take up water.

Refined by @blackman2016, incorporating more well-known hydraulics traits into the framework.

Discuss Martin-StPaul (2017); key role of minimum conductance in a model of desiccation dynamics

Various new models that aim to simulate desiccation have a g~min~ term in it, but it is not yet well constrained by data.

# VI. Cuticular conductance

@@

\
Although we use the term cuticular conductance freely, this transport pathway does not represent a true conductance since water does not diffuse as a gas through the cuticle. Instead, it dissolves into the medium of the cuticle, diffuses through the solid matrix, and is desorped at the outer edge of the cuticle [@kerstiens1996; @riederer2001; @schreiber1996]. The main barrier to diffusion is actually a very thin layer at the leaf surface. Because most of the resistance is located in such a thin layer, g~cut~ does not correlate with the thickness of the cuticle  [@priestley1943; @riederer2001; @anfodillo2002; @schuster2016b]. Nonetheless there is considerable variation in cuticle thickness between plant species [@schuster2016a], as well as within species in response to drought (cite), along altitudinal transects [@delucia1984], and even with increasing height in the canopy of very tall trees [@woodruff2010]. If the cuticle thickness does not directly affect its conductance, what is the value of increasing the cuticle?

\
It is important to bear in mind other functions of the cuticle, including attenuation of radiation, as a barrier for fungal pathogens, and various interactions with insects [@mueller2008] [@kerstiens1996c;@riederer2008]. The thickness of the cuticle per se may also have to do with mechanical strength, as @onoda2012 reported that thicker cuticles are more resistant to tearing (the force to tear and tensile strength both approx proportional to cuticle thickness). Thick cuticles may also be useful because  exposure to wind has been shown to dislodge cuticular compounds by absrasion [@rogge1993], thereby dramatically increasing the conductance of the cuticle [@hadley1983;@grace1974;@vangardingen1991]. Similarly, exposure to simulated rain damages the cuticle and increases its conductance [@baker1986]. Such exposure to the elements may be responsible for the typical gradual increase in g~min~ with leaf age in field-grown plants [@jordan2007].

\
Despite considerable work on the topic, no clear relationship between chemical composition or structure of the cuticle and its conductance has emerged. Recently, @schuster2016a presented a comprehensive study of cuticle chemical composition, but this variation could not explain variation in g~cut~. Similarly, @hauke1998 found no relationship between the gradual decrease in g~cut~ in *Hedera helix* during leaf maturation and cuticle wax amount, mean chain lengths, or cuticle weight.

\
Water travels more freely across the cuticle than CO2 [@boyer1997; @boyer2015]. This observation was used by @manzoni2011 to recalculate leaf water-use efficiency for droughted plants, but we caution against this because g~cut~ is only one component of g~min~, and CO~2~ will travel through leaky and incompletely closed stomata. Nonetheless it is an interesting observation that could have major consequences for modelling and interpreting water-use efficiency data during drought conditions.

# VII. Nighttime conductance

@@

# VIII. Environmental and ecological variation in minimum conductance

## 1. Contribution of stomata to minimum conductance

@@

## 2. Minimum conductance shows little meaningful interspecific variation

@@

@@

Comparing all available data, we did not find significant relationships between rainfall, aridity, or temperature and g~min~ (see SuppInfo 2).

## 3. Acclimation to drought

@@

The second case study was on *Hakea* plants. Details here (measurement T?).

@@

## 4. Acclimation and instantaneous response to temperature

@@

Lower g~min~ result is consistent with @duarte2016, who reported lower g~0~ and g~dark~ in a heat wave treatment (Douglas-fir), which persisted for some time. This is not due to a direct effect of temperature on in cuticular conductance, which increases with temperature (Schreiber 2001).

A case study on *Eucalyptus parramattensis* growing in twelve whole-tree chambers (see Drake for methods), either tracking ambient conditions or exposed to a constant +3 C air temperature. Our results showed a large  acclimation effect but no evidence for a direct T effect (over the narrow range in T compared to Schreiber).

## 5. Other factors

Also discussed above in Section~(cuticular conductance), the cuticle is constantly exposed to physical (wind, rain) and biotic (insects, fungi) agents which damage the cuticle over time.

--> more in database

@@

# Appendix or SuppInfo

\
To illustrate the effect of g~0~ on the modelled response of g~s~ to PPFD and T~leaf~, it is necessary to embed the stomatal conductance model in a coupled leaf gas exchange model, which is formed by three equations; Eq. 1 giving the dependence of g~s~ on A~n~ and some environmental drivers, the diffusion constraint,
